\documentclass[conference]{IEEEtran}
%\documentclass[conference, onecolumn]{IEEEtran}
%\documentclass[10pt,twocolumn]{article} 
%\usepackage{latex8}
%\usepackage{times}

%\renewcommand\thesection{\arabic{section}}
%\renewcommand\thesubsection{\thesection.\arabic{subsection}}
%\renewcommand\thesubsubsection{\thesubsection.\arabic{subsubsection}}

%\renewcommand\thesectiondis{\arabic{section}}
%\renewcommand\thesubsectiondis{\thesectiondis.\arabic{subsection}}
%\renewcommand\thesubsubsectiondis{\thesubsectiondis.\arabic{subsubsection}}



\usepackage[pdftex]{graphicx}
\usepackage{epsfig}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{subfig}
\usepackage{cite}
\usepackage{paralist}
\usepackage{tikz}

%\usepackage{etoolbox}

\usepackage{color}

\newcommand{\ceiling}[1]{\left\lceil{#1}\right\rceil}
\newcommand{\floor}[1]{\left\lfloor{#1}\right\rfloor}
\newcommand{\setof}[1]{\left\{{#1}\right\}}
\newcommand{\set}[2]{\{{#1}\mid{#2}\}}
\newcommand{\equals}{\stackrel{\mathrm{def}}{=}}

\newcommand{\jj}[1]{\textcolor{blue}{jj: #1 : endjj}}
\newcommand{\kh}[1]{\textcolor{magenta}{kevin: #1 : endkevin}}
\newcommand{\framework}[1]{$\mathbf{k^2U}$}
\newcommand{\frameworkkq}[1]{$\mathbf{k^2Q}$}
\newcommand{\Cong}[1]{\textcolor{red}{cong: #1 : endcong}}


%%%%%%%%%%%%%%%
 \def\myendproof{\qed}%{\ \vbox{\hrule\hbox{%
   %\vrule height1.3ex\hskip0.8ex\vrule}\hrule }}\par}
 %\renewenvironment{proof}{\noindent{\bf Proof. }}{\myendproof}
%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%
 \newenvironment{appProof}[1]{\noindent{\bf Proof of
     #1. }}{\myendproof\vskip 0.1in}
 %%%%%%%%%%%%%%

\newtheorem{theorem}{Theorem}
\newtheorem{Lemma}{Lemma}
\newtheorem{Corollary}{Corollary}
\newtheorem{example}{Example}
\newtheorem{claim}{Claim}
\newtheorem{definition}{Definition}
\newtheorem{rrule}{Rule}
\newtheorem{remark}{Remark}
\newtheorem{observation}{Observation}
\newtheorem{Property}{Property}
\addtolength{\textheight}{8pt}

\usetikzlibrary{%
  arrows,%
  shapes.misc,% wg. rounded rectangle
  shapes.arrows,%
  chains,%
  matrix,%
  positioning,% wg. " of "
  scopes,%
  decorations.pathmorphing,% /pgf/decoration/random steps | erste Graphik
  shadows%
}

\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}
\tikzstyle{materia}=[draw, fill=white, text width=1.0em, text centered,
  minimum height=4em,drop shadow]
\tikzstyle{practica} = [materia, text width=18em, minimum width=8em,
align =left,
  minimum height=3em, rounded corners, drop shadow]
\tikzstyle{texto} = [above, text width=6em, text centered]
\tikzstyle{linepart} = [draw, thick, color=blue!50, -latex', dashed]
\tikzstyle{line} = [draw, line width = 2pt, color=blue!50, -latex']
\tikzstyle{ur}=[draw, text centered, minimum height=0.01em]

\bibliographystyle{abbrv} 

%\newif\ifpaper
%\paperfalse

\pagestyle{plain}

\begin{document}
  
\title{A Unifying Response Time Analysis Framework for Dynamic Self-Suspending Tasks}

\author{Jian-Jia Chen\IEEEauthorrefmark{1}, Geoffrey Nelissen\IEEEauthorrefmark{4}, Wen-Hung Huang\IEEEauthorrefmark{1}\\
\IEEEauthorrefmark{1} TU Dortmund University, Germany\\
\IEEEauthorrefmark{4} CISTER/INESC-TEC, ISEP, Polytechnic Institute of Porto, Portugal \\
Emails: \{jian-jia.chen, wen-hung.huang\}@tu-dortmund.de, grrpn@isep.ipp.pt
}

\maketitle

\begin{abstract}
  For real-time embedded systems, self-suspending behaviors can cause
  substantial performance/schedulability degradations. In this paper,
  we focus on preemptive fixed-priority scheduling for the dynamic
  self-suspension task model on uniprocessor. This
  model assumes that a job of a task can dynamically suspend itself during its execution (for instance, to wait for shared resources or access co-processors or external devices).
  The total suspension time of a job is upper-bounded, but this dynamic behavior drastically influence the interference generated on lower-priority tasks. The state-of-the-art results for this task model can be classified
  into three categories (i) modeling suspension as computation, (ii)
  modeling suspension as release jitter, and (iii) modeling suspension as a blocking term.
  However, several results associated to the release jitter approach were recently proven to be erroneous, and the concept of modeling suspension as blocking was never
  formally proven correct. This paper presents a unifying
  response time analysis framework for the dynamic self-suspending
  task model. We provide a rigorous proof and show that the existing analyses pertaining to the three categories mentioned above are analytically dominated by our proposed solution. Therefore, all those techniques are in fact correct, although
  inferior to the proposed response time analysis in this paper. The
  evaluation results show that our analysis framework can generate huge
  improvements (an increase of up to $50\%$ of the number of task sets
  deemed schedulable) over these state-of-the-art analyses.
\end{abstract}

\input{introduction}

\input{model}

\input{background}

\input{rationale}

\input{analysis}

\input{dominance}

\input{approximation}













\input{experiments}

\section{Conclusion}
\label{sec:conclusion}


In this paper, we studied the preemptive fixed-priority scheduling of
dynamic self-suspending tasks running on a uniprocessor platform. This
paper presents a unifying response time analysis framework in
Theorem~\ref{theorem:general-framework} and
Corollary~\ref{corollary:general-framework}. We show that this result
analytically dominates all the existing analyses presented in
Section~\ref{sec:existing-analyses}, and, by doing such, we also implicitly
proved the correctness of all these analyses. Although
Corollary~\ref{corollary:general-framework} requires exponential time
complexity, we show that a simpler algorithm presented in Section~\ref{sec:linear-approximation} can help accelerate the analysis while outputting good results.

%In fact, the response time analysis and the schedulability tests in
%Section~\ref{sec:existing-analyses} and
%Theorem~\ref{theorem:general-framework} can be further approximated to
%polynomial-time complexity. For example, by over-approximating the
%jitter used in Section~\ref{sec:jitter}, Liu and Chen
%in \cite{LiuChen:rtss2014} provided a utilization-based schedulability
%test based on a hyperbolic-form. Moreover, as presented in
%Section~\ref{sec:suspension-blocking}, Liu in \cite{Liu:2000:RS:518501}
%in fact used the blocking time for a schedulability test based on the
%RM utilization bound. Our test in
%Theorem~\ref{theorem:linear-time-test} is a linear-time test. The
%above tests can be further improved by using the k2U and the k2Q
%frameworks
%in \cite{ChenHLRTSS2015,DBLP:journals/corr/ChenHL15b}. However, as
%polynomial-time schedulability tests are not the focus of this paper,
%such options were not seriously discussed here.

\noindent{\bf Acknowledgements.} This paper is supported by DFG, as part of the Collaborative Research Center SFB876 (http://sfb876.tu-dortmund.de/). This work was also partially supported by National Funds through FCT/MEC (Portuguese Foundation for Science and Technology) and co-financed by ERDF (European Regional Development Fund) under the PT2020 Partnership, within project UID/CEC/04234/2013 (CISTER); also by FCT/MEC and the EU ARTEMIS JU within project(s) ARTEMIS/0003/2012 - JU grant nr. 333053 (CONCERTO) and ARTEMIS/0001/2013 - JU grant nr. 621429 (EMC2).

\bibliography{../bibliography/biblio-summary,../bibliography/biblio}{}

\end{document}
